Data Flow Analysis\\in the Presence of Correlated Calls

### 2014-12-05 content/background.tex
add comments on what to remove

\subsection{Related Work}
IFDS is a version of the functional approach to data-flow analysis developed by M.\,Sharir and A.\,Pnueli~\cite{pnueli1981two}. Their algorithm is based on computing \textit{summary functions} that return the data-flow value at the end of a procedure, given the data-flow value at the start of the procedure. IFDS problems form a more restricted set of data-flow problems: unlike in the functional approach, IFDS flow functions have to be distributive, and the set of data-flow facts $D$ has to be finite. However, the IFDS algorithm is more general than Sharir's and Pnueli's algorithm in that it can handle programs containing local variables and parameters in recursive methods.

@@

\begin{example}
  Consider the program in Listing~\ref{list:examplejava}. The supergraph corresponding to that program is shown in Figure~\ref{fig:examplesupergraph}.
\begin{figure}
  \centering
  \begin{minipage}{\textwidth}
    \inputMinted{java}{examplesupergraph.java}
  \end{minipage}
  \caption{An example Java program}
  \label{list:examplejava}
\end{figure}

>>>>

\subsection{Related Work}
\otodo{should we move Related Work to near the end of the paper?}
IFDS is a version of the functional approach to data-flow analysis developed by M.\,Sharir and A.\,Pnueli~\cite{pnueli1981two}. Their algorithm is based on computing \textit{summary functions} that return the data-flow value at the end of a procedure, given the data-flow value at the start of the procedure. IFDS problems form a more restricted set of data-flow problems: unlike in the functional approach, IFDS flow functions have to be distributive, and the set of data-flow facts $D$ has to be finite. However, the IFDS algorithm is more general than Sharir's and Pnueli's algorithm in that it can handle programs containing local variables and parameters in recursive methods.

@@

\begin{example}
  Consider the program in Listing~\ref{list:examplejava}. The supergraph corresponding to that program is shown in Figure~\ref{fig:examplesupergraph}.
  \otodo{Can we merge the examples from Figures 1 and 2 into a single example program?}
\otodo{I think we can remove the supergraph Figure 3.}
\begin{figure}
  \centering
  \begin{minipage}{\textwidth}
    \inputMinted{java}{examplesupergraph.java}
  \end{minipage}
  \caption{An example Java program}
  \label{list:examplejava}
\end{figure}

### 2014-12-05 content/eval.tex
add comments on what to remove

\subsubsection{IFDS}
As described in Section~\ref{sec:ifdsdef}, an IFDS problem is defined in terms of an exploded supergraph. The control-flow supergraph of an input program can be retrieved using WALA. Hence, our implementation of an IFDS problem should be able to convert a control-flow supergraph into an exploded supergraph.

We represent an IFDS problem with a trait, or protocol, that contains declarations of four \textit{flow functions}. Each function has type 
\[
  F:\,(N\times D\times N)\to2^D
\]
and defines a set of edges on the exploded graph. 
Given an edge $(n_1,\,n_2)$ of the control-flow supergraph and the fact $d_1$ that 
corresponds to the source node $n_1$, $F(n_1,\,d_1,\,n_2)$ returns the set of all facts $d_2\in D_2$ such that $((n_1,\,d_1),\,(n_2,\,d_2))\in E^\#$\footnote{In each invocation of a flow function, the fact $d_1$ is provided by the IDE algorithm.}.
The four functions are:
\begin{itemize}
  \item \textsf{call-start}, for inter-procedural edges from a call node to the start node of the target method;
  \item \textsf{call-return}, for intra-procedural edges from a call node to its return node;
  \item \textsf{end-return}, for inter-procedural edges from the end node of a method to the return node of the callee;
  \item \textsf{default}, for all other intra-procedural edges.
\end{itemize}

@@

\begin{example}
  Consider the supergraph in Figure~\ref{fig:exampleexploded}. The call-to-start flow function from method \verb'main' to \verb'f' looks as follows:
  \begin{align*}
    \textsf{call-start}(\highlight{\textsf{call}_{\texttt{A.f}}}{greyblue},\,\texttt a,\,\highlight{\textsf{start}_\texttt{f}}{lightsalmonpink})
    &=r_\texttt{main}(\texttt a)\cup\{\texttt s\}\\
    &=\{\texttt s\}.
  \end{align*}
  We can see that correspondingly, the exploded supergraph contains an edge from $(\highlight{\textsf{call}_{\texttt{A.f}}}{greyblue},\,\texttt a)$ to $(\highlight{\textsf{start}_\texttt{f}}{lightsalmonpink},\,\texttt s)$.
\end{example}

@@

\subsubsection{Conversion from IFDS to IDE}
We implemented the equivalence transformation $\transEq$ and the correlated-calls transformation $\transCC_\rcc$ from IFDS to IDE described in Section~\ref{sec:equivtrans}. To run an IFDS problem, we converted it to an IDE problem using $\transEq$ and $\transCC_\rcc$ and used our IDE analysis algorithm to run the latter.

@@

We also implemented a correlated-call transformation from IFDS into IDE problems that consider correlated calls. This transformation is described in Section~\ref{sec:cctrans}.

@@

\subsubsection{Benchmark Testing}
To assess the benefit of the correlated-calls analysis, we counted the frequencies of correlated-call occurrences in the Dacapo benchmarks~\cite{blackburn2006dacapo}. We then ran the normal- and correlated-call-taint analysis on the Dacapo benchmarks to see what improvement we would get from the correlated-calls analysis.

@@

\subsection{Future Work}
In this section we point out the limitations of the correlated-calls analysis and suggest improvements to the analysis for future work.

@@

>>>>

\begin{odelete}
\subsubsection{IFDS}
\otodo{We should incorporate the four functions into the explanation of IFDS in the background. Then we don't need them here.}
As described in Section~\ref{sec:ifdsdef}, an IFDS problem is defined in terms of an exploded supergraph. The control-flow supergraph of an input program can be retrieved using WALA. Hence, our implementation of an IFDS problem should be able to convert a control-flow supergraph into an exploded supergraph.

We represent an IFDS problem with a trait, or protocol, that contains declarations of four \textit{flow functions}. Each function has type 
\[
  F:\,(N\times D\times N)\to2^D
\]
and defines a set of edges on the exploded graph. 
Given an edge $(n_1,\,n_2)$ of the control-flow supergraph and the fact $d_1$ that 
corresponds to the source node $n_1$, $F(n_1,\,d_1,\,n_2)$ returns the set of all facts $d_2\in D_2$ such that $((n_1,\,d_1),\,(n_2,\,d_2))\in E^\#$\footnote{In each invocation of a flow function, the fact $d_1$ is provided by the IDE algorithm.}.
The four functions are:
\begin{itemize}
  \item \textsf{call-start}, for inter-procedural edges from a call node to the start node of the target method;
  \item \textsf{call-return}, for intra-procedural edges from a call node to its return node;
  \item \textsf{end-return}, for inter-procedural edges from the end node of a method to the return node of the callee;
  \item \textsf{default}, for all other intra-procedural edges.
\end{itemize}
\end{odelete}

@@

\begin{odelete}
\begin{example}
  Consider the supergraph in Figure~\ref{fig:exampleexploded}. The call-to-start flow function from method \verb'main' to \verb'f' looks as follows:
  \begin{align*}
    \textsf{call-start}(\highlight{\textsf{call}_{\texttt{A.f}}}{greyblue},\,\texttt a,\,\highlight{\textsf{start}_\texttt{f}}{lightsalmonpink})
    &=r_\texttt{main}(\texttt a)\cup\{\texttt s\}\\
    &=\{\texttt s\}.
  \end{align*}
  We can see that correspondingly, the exploded supergraph contains an edge from $(\highlight{\textsf{call}_{\texttt{A.f}}}{greyblue},\,\texttt a)$ to $(\highlight{\textsf{start}_\texttt{f}}{lightsalmonpink},\,\texttt s)$.
\end{example}
\end{odelete}

@@

\begin{odelete}
\subsubsection{Conversion from IFDS to IDE}
We implemented the equivalence transformation $\transEq$ and the correlated-calls transformation $\transCC_\rcc$ from IFDS to IDE described in Section~\ref{sec:equivtrans}. To run an IFDS problem, we converted it to an IDE problem using $\transEq$ and $\transCC_\rcc$ and used our IDE analysis algorithm to run the latter.

@@

We also implemented a correlated-call transformation from IFDS into IDE problems that consider correlated calls. This transformation is described in Section~\ref{sec:cctrans}.
\end{odelete}

@@

\subsubsection{Benchmark Testing}
\otodo{We need to decide what exactly we want to include in the evaluation.
    The numbers of correlated calls are interesting. But I think we want
    to leave out all mention of the taint analysis. Perhaps we should
    have no empirical results at all (and only the theory/proofs)?}
To assess the benefit of the correlated-calls analysis, we counted the frequencies of correlated-call occurrences in the Dacapo benchmarks~\cite{blackburn2006dacapo}. We then ran the normal- and correlated-call-taint analysis on the Dacapo benchmarks to see what improvement we would get from the correlated-calls analysis.

@@

\begin{odelete}
\subsection{Future Work}
\otodo{We should convert this section into two or three sentences to be added to the conclusion. They should focus especially on the interprocedurally-correlated receivers.}
In this section we point out the limitations of the correlated-calls analysis and suggest improvements to the analysis for future work.

@@

\end{odelete}

### 2014-12-05 content/intro.tex
add comments on what to remove

\section{Introduction}
Static program analysis aims to discover properties of computer programs without running them.
Static analysis has applications in compiler optimization, development of programming tools, and computer security, among others.
As an example, we might want to analyze a program to know which variables are constants. 
We could then write a compiler optimization that ensures that the values of those variables are computed only once.
Alternatively, we could use the information about constant variables in an integrated development environment; for instance, to notify the user when an if-expression executes only one of its branches because its test condition has a constant value.

@@

We improve the precision of IFDS problem solutions by eliminating infeasible paths that occur in object-oriented programs in the presence of \textit{correlated method calls}~--- polymorphic calls that are invoked on the same object~\cite{DBLP:journals/scp/Tip15}.

@@

\subsection{IFDS and IDE}
The IFDS framework is a precise and efficient algorithm for data-flow analysis. IFDS was developed in 1995 by T.\,Reps, S.\,Horwitz, and M.\,Sagiv at the University of Wisconsin and has been used to solve a variety of data-flow analysis problems~\cite{bodden2013spl,naeem2008typestate,DBLP:conf/birthday/KreikerRRSWY13,tripp2009taj}. The IFDS analysis is a version of the classic \textit{functional approach} to data-flow analysis proposed by M.\,Sharir and A.\,Pnueli~\cite{pnueli1981two}.

>>>>

\section{Introduction}
\otodo{This section should be shrunk a lot (not completely deleted)}
\begin{odelete}
Static program analysis aims to discover properties of computer programs without running them.
Static analysis has applications in compiler optimization, development of programming tools, and computer security, among others.
As an example, we might want to analyze a program to know which variables are constants. 
We could then write a compiler optimization that ensures that the values of those variables are computed only once.
Alternatively, we could use the information about constant variables in an integrated development environment; for instance, to notify the user when an if-expression executes only one of its branches because its test condition has a constant value.

@@

We improve the precision of IFDS problem solutions by eliminating infeasible paths that occur in object-oriented programs in the presence of \textit{correlated method calls}~--- polymorphic calls that are invoked on the same object~\cite{DBLP:journals/scp/Tip15}.
\end{odelete}

@@

\subsection{IFDS and IDE}
\otodo{This section should be moved to the Background section (and suitably merged with the text there).}
The IFDS framework is a precise and efficient algorithm for data-flow analysis. IFDS was developed in 1995 by T.\,Reps, S.\,Horwitz, and M.\,Sagiv at the University of Wisconsin and has been used to solve a variety of data-flow analysis problems~\cite{bodden2013spl,naeem2008typestate,DBLP:conf/birthday/KreikerRRSWY13,tripp2009taj}. The IFDS analysis is a version of the classic \textit{functional approach} to data-flow analysis proposed by M.\,Sharir and A.\,Pnueli~\cite{pnueli1981two}.
